Adaptive group-regularized logistic elastic net regression

### 2019-03-19 docs/manuscript.Rnw
fixed all cross-document references

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
titlesec,multicol,hyperref,verbatim,threeparttable,booktabs,dsfont,parskip}

% Add history information for the article if required
%\history{Received August 1, 2010;
%	revised October 1, 2010;
%	accepted for publication November 1, 2010}

@@

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features 
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, (b) a summary of prior 
		information, and (c) omics annotation. The inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward in the standard regression setting. As a solution
		to this problem, we propose a group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed 
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an 
		approximate empirical-variational Bayes framework. Simulations and an
		application to a colon cancer microRNA study show that, if the partitioning
		of the features is informative, classification performance and feature
		selection are indeed enhanced.}
	\end{abstract}

@@

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see Supplementary Material (SM) Section [! section reference here] for 
	details on this prior) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see Section [! section ref here] 
	of the SM for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
	respectively.

	To simplify the computations of our variational approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
	calculus of variations gives the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions in our variational Bayes implementation for the model parameters
	as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section 5 for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 6 of the SM we show that 
	informed calculation of the parameters results in a significant reduction of 
	computational complexity.

@@

	To estimate the penalty multipliers, the intractable posterior expectation
	in (\ref{eq:ebEM}) is approximated using the variational posterior:
	$$
	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \]
	= \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
	\lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	$$
	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full 
	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
	of the new penalty multipliers is now given by:
	\begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}
	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
	closed form, this convex problem is easily solved by a numerical optimisation 
	routine. The full procedure is summarized in Section 8 of the SM.

@@

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section [! section ref here] 
	of 
	the SM. In the first scenario, the distribution of model parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In the second scenario, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate, while we expect the other methods to perform better, due to a
	lack of overfitting. In scenario three, we expect all methods to underperform
	due to the high correlations between the features. The fourth 
	scenario is conjectured to yield a better \texttt{gren} performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	incorporate a differential signal between
	the groups. Scenario five is expected to yield a bad performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

@@

	[! results here]

	Additional results and performance measures are presented in Section 
	[! section ref here] of the SM.

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
  three methods, we only present \texttt{cMCP} here; the results for 
  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
  [! section ref here]. For the methods with a
  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
  performance and refer
  the reader to SM Section [! section ref here] for comparisons with 
  different $\alpha$'s.

  To estimate performance, we split the data into 61 training and 27 test 
  instances, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present the AUC for a
  range of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
  performance measures are presented in Section [! section ref here]
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation on this dataset in SM Section 
  [! section ref here]. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global $\lambda$ on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
  all group lasso extensions that we investigated.
  (SM [! section ref here]). Overall, it seems as though the inclusion of the
  extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

@@

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used an extension of
  \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarray expression levels, normalised to account
  for a difference in scale compared to the RNAseq training data. We present the
  results on this validation set, together with the estimated penalty 
  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
  comparisons and performance measures are presented in the SM Section 
  [! section ref here].

@@

	Although our method can be considered weakly adaptive and is essentially 
	a reweighted elastic net, it is different from 
	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. As can be seen from both the simulations in 
	Section \ref{sec:simulations} and the applications to different outcomes 
	(prognostic or diagnostic) and marker types 
	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
	\ref{sec:lymphnode}, and SM Sections [! section refs here], adaptation to 
	external data may be beneficial for prediction and feature selection. 
	We believe that this is due to the `borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better estimates.

	A common criticism of the lasso (and elastic net) is its instability of 
	feature 
	selection: different data instances lead to different sets of selected
	features. We investigated the stability of selection on the data from Section
	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
	the original data and calculated the sizes of all intersections of selected
	features. We compared \texttt{gren} to the 
	regular elastic net and found that the inclusion of extra information
	increased the selection stability (see SM Section [! section ref here]). 
	In addition, we investigated 
	the variance and bias in penalty multiplier estimation by examining the
	distribution
	of estimated penalty parameters from 100 random partitionings of the features
  (see SM Section [! section ref here]). We found that the penalty parameters
  tend to cluster around one, as expected for an unbiased estimation method.

	As touched upon in Section \ref{sec:introduction}, another obvious 
	comparison is to the group lasso \cite[]{meier_group_2008,
	yuan_model_2006} and its extensions. Although it is similar in the sense that 
	it shrinks on the 
	group level, the group lasso is built upon an entirely different philosophy: 
	its intended application is to small interpretable groups of features, like, 
	for example, dummies of a categorical variable. Another difference between 
	\texttt{gren} and the group lasso is the number of penalty parameters. 
	\texttt{gren} estimates one parameter per group, while the group lasso
	estimates one overall penalty parameter; it is thereby less flexible in 
	differential shrinkage of the parameters. In addition, in the simulations and
	data applications of \cite{van_de_wiel_better_2016}, group lasso prediction 
	performed inferior to \texttt{GRridge} prediction.

>>>>

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
xr-hyper,hyperref}
\externaldocument[sm-]{supplement}

@@

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features 
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, (b) a summary of prior 
		information, and (c) omics annotation. The inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward in the standard regression setting. As a solution
		to this problem, we propose a group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed 
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an 
		approximate empirical-variational Bayes framework. Simulations and
		applications in two different molecular marker and cancer type studies show
		that, if the partitioning of the features is informative, classification 
		performance and feature
		selection are indeed enhanced.}
	\end{abstract}

@@

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
	details on this prior) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see Section 
	\ref{sm-sec:objectivefunction} of the SM for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
	respectively.

	To simplify the computations of our variational approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
	\ref{sm-sec:polyagamma} of the SM for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
	calculus of variations gives the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions in our variational Bayes implementation for the model parameters
	as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
	for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 
	\ref{sm-sec:computationalcomplexity} of the SM we show that 
	informed calculation of the parameters results in a significant reduction of 
	computational complexity.

@@

	To estimate the penalty multipliers, the intractable posterior expectation
	in (\ref{eq:ebEM}) is approximated using the variational posterior:
	\begin{equation}\label{eq:approximatedloglikelihood}
	  \E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | 
	  \blambda'^{(k)} \]
	  = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
	  \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	\end{equation}
	where $C$ is constant in $\blambda'$ (see SM Section 
	\ref{sm-sec:expectedloglikelihood} for the full 
	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
	of the new penalty multipliers is now given by:
	\begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}
	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
	closed form, this convex problem is easily solved by a numerical optimisation 
	routine. The full procedure is summarized in Section 
	\ref{sm-sec:fullestimationprocedure} of the SM.

@@

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 
	\ref{sm-sec:simulations} of 
	the SM. In the first scenario, the distribution of model parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In the second scenario, which is similar to
	the simulations presented in the original group lasso paper 
	\cite[]{yuan_model_2006}, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate, while we expect the other methods to perform better, due to a
	lack of overfitting. In scenario three, we expect all methods to underperform
	due to the high correlations between the features. The fourth 
	scenario is conjectured to yield a better \texttt{gren} performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	incorporate a differential signal between
	the groups. Scenario five is expected to yield a bad performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

@@

<<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
@

	Additional results and performance measures are presented in Section 
	\ref{sm-sec:simulations} of the SM.

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
  three methods, we only present \texttt{cMCP} here; the results for 
  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
  [! section ref here]. For the methods with a
  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
  performance and refer
  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
  different $\alpha$'s.

  To estimate performance, we split the data into 61 training and 27 test 
  instances, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present the AUC for a
  range of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
  performance measures are presented in Section \ref{sm-sec:colorectal}
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation on this dataset in SM Section 
  \ref{sm-sec:colorectal}. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global $\lambda$ on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
  all group lasso extensions that we investigated.
  (SM Section \ref{sm-sec:colorectal}). Overall, it seems as though the 
  inclusion of the extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

@@

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used an extension of
  \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarray expression levels, normalised to account
  for a difference in scale compared to the RNAseq training data. We present the
  results on this validation set, together with the estimated penalty 
  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
  comparisons and performance measures are presented in the SM Section 
  \ref{sm-sec:lymphnode}.

@@

	Although our method can be considered weakly adaptive and is essentially 
	a reweighted elastic net, it is different from 
	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. As can be seen from both the simulations in 
	Section \ref{sec:simulations} and the applications to different outcomes 
	(prognostic or diagnostic) and marker types 
	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
	\ref{sm-sec:cervical}, adaptation to 
	external data may be beneficial for prediction and feature selection. 
	We believe that this is due to the `borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better estimates.

	A common criticism of the lasso (and elastic net) is its instability of 
	feature 
	selection: different data instances lead to different sets of selected
	features. We investigated the stability of selection on the data from Section
	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
	the original data and calculated the sizes of all intersections of selected
	features. We compared \texttt{gren} to the 
	regular elastic net and found that the inclusion of extra information
	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
	In addition, we investigated 
	the variance and bias in penalty multiplier estimation by examining the
	distribution
	of estimated penalty parameters from 100 random partitionings of the features
  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
  tend to cluster around one, as expected for an unbiased estimation method.

	As touched upon in Section \ref{sec:introduction}, another obvious 
	comparison is to the group lasso \cite[]{meier_group_2008,
	yuan_model_2006} and its extensions. Although the group lasso is similar in 
	the sense that it shrinks on the 
	group level, it is built upon an entirely different philosophy: 
	its intended application is to small interpretable groups of features, like, 
	for example, dummies of a categorical variable. Another difference between 
	\texttt{gren} and the group lasso is the number of penalty parameters. 
	\texttt{gren} estimates one parameter per group, while the group lasso
	estimates one overall penalty parameter; it is thereby less flexible in 
	differential shrinkage of the parameters. In addition, we show in our 
	simulations and data applications that \texttt{gren} is competitive and often
	even superior to (extensions of) the group lasso.
